<!doctype html>

<html class="no-js" lang="en">

<head>


	<!-- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

	DIY.transcriptomics

	Personal Theme by https://jekyllthemes.io
	Premium + free Jekyll themes for your blog or website.

	- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -->


	<meta charset="utf-8">
	<meta http-equiv="X-UA-Compatible" content="IE=edge,chrome=1">

	<!-- Page Info -->
	<link rel="shortcut icon" href="/images/favicon.png">
	<title>DIYtranscriptomics – DIY.transcriptomics</title>
	<meta name="description" content="">

	<!-- Twitter Card -->
	<meta name="twitter:card" content="summary_large_image">
	<meta name="twitter:title" content="DIYtranscriptomics – DIY.transcriptomics">
	<meta name="twitter:description" content="">
	<meta name="twitter:image:src" content="http://diytranscriptomics.com">

	<!-- Facebook OpenGraph -->
	<meta property="og:title" content="DIYtranscriptomics – DIY.transcriptomics" />
	<meta property="og:description" content="" />
	<meta property="og:image" content="http://diytranscriptomics.com" />

	
	<!-- Font Embed Code -->
	<link href="https://fonts.googleapis.com/css?family=Muli:300,400,600,700" rel="stylesheet">
	

	<!-- Styles -->
	<meta name="viewport" content="width=device-width, initial-scale=1">
	<link rel="stylesheet" href="/css/style.css">
	
	<!-- Icons -->
	<script defer src="https://use.fontawesome.com/releases/v5.1.1/js/solid.js" integrity="sha384-GXi56ipjsBwAe6v5X4xSrVNXGOmpdJYZEEh/0/GqJ3JTHsfDsF8v0YQvZCJYAiGu" crossorigin="anonymous"></script>
	<script defer src="https://use.fontawesome.com/releases/v5.1.1/js/brands.js" integrity="sha384-0inRy4HkP0hJ038ZyfQ4vLl+F4POKbqnaUB6ewmU4dWP0ki8Q27A0VFiVRIpscvL" crossorigin="anonymous"></script>
	<script defer src="https://use.fontawesome.com/releases/v5.1.1/js/fontawesome.js" integrity="sha384-NY6PHjYLP2f+gL3uaVfqUZImmw71ArL9+Roi9o+I4+RBqArA2CfW1sJ1wkABFfPe" crossorigin="anonymous"></script>

	
	<!-- Custom Styles -->
	<style></style>
	

	
	<!-- Analytics Code -->
	
	

	
	<!-- Extra Header JS Code -->
	
	
	
</head>


<body class="loading ajax-loading" data-site-url="http://diytranscriptomics.com" data-page-url="/Code/files/flexdashboard_essentials.Rmd">


	<header class="header">

	<div class="wrap">

		
		<a href="/" class="header__title">
			DIY.transcriptomics
		</a>
		

		<div class="menu">
			<div class="menu__toggle js-menu-toggle">
				<div class="menu__toggle__icon"><span></span></div>
			</div>
			<div class="menu__wrap">
				<ul class="menu__list">
					
					<li class="menu__list__item">
						<a href="/" class="menu__list__item__link">Syllabus</a>
					</li>
					
					<li class="menu__list__item">
						<a href="/about" class="menu__list__item__link">About</a>
					</li>
					
					<li class="menu__list__item">
						<a href="/data" class="menu__list__item__link">Data</a>
					</li>
					
					<li class="menu__list__item">
						<a href="/scripts" class="menu__list__item__link">Scripts</a>
					</li>
					
					<li class="menu__list__item">
						<a href="/help" class="menu__list__item__link">Help</a>
					</li>
					
					<li class="menu__list__item">
						<a href="/register" class="menu__list__item__link">Register</a>
					</li>
					
				</ul>
			</div>
		</div>

	</div>

</header>


	<div class="loader"><svg width="120" height="30" viewBox="0 0 120 30" xmlns="http://www.w3.org/2000/svg"><circle cx="15" cy="15" r="15"><animate attributeName="r" from="15" to="15" begin="0s" dur="0.8s" values="15;9;15" calcMode="linear" repeatCount="indefinite" /><animate attributeName="fill-opacity" from="1" to="1" begin="0s" dur="0.8s" values="1;.5;1" calcMode="linear" repeatCount="indefinite" /></circle><circle cx="60" cy="15" r="9" fill-opacity="0.3"><animate attributeName="r" from="9" to="9" begin="0s" dur="0.8s" values="9;15;9" calcMode="linear" repeatCount="indefinite" /><animate attributeName="fill-opacity" from="0.5" to="0.5" begin="0s" dur="0.8s" values=".5;1;.5" calcMode="linear" repeatCount="indefinite" /></circle><circle cx="105" cy="15" r="15"><animate attributeName="r" from="15" to="15" begin="0s" dur="0.8s" values="15;9;15" calcMode="linear" repeatCount="indefinite" /><animate attributeName="fill-opacity" from="1" to="1" begin="0s" dur="0.8s" values="1;.5;1" calcMode="linear" repeatCount="indefinite" /></circle></svg></div>

	<div class="page-loader"></div>

	
	<div class="page">

		<div class="page__content" data-page-title="DIYtranscriptomics – DIY.transcriptomics">

			<section class="hero hero--single">

	<div class="hero__image" style="background-image: url()">
		<div class="hero__overlay"></div>
	</div>

	<div class="wrap">

		<h1>DIYtranscriptomics</h1>
		<p></p>

	</div>

</section>

<section class="single">

	<div class="wrap">

		<article class="single-post">

			```{r setup, include=FALSE}
library(tidyverse)
library(tximport)
library(biomaRt)
library(RColorBrewer) 
library(reshape2) 
library(genefilter)
library(edgeR) 
library(matrixStats)
library(hrbrthemes)
library(gplots) 
library(limma)
library(heatmaply)
library(d3heatmap)
library(DT)
library(gt)
library(plotly)
knitr::opts_chunk$set(message = FALSE)
```

Overview {data-icon="fa-signal"}
===================================== 

Row {data-height=400}
------------------------------------------------------------------------------

### Signal distribution: raw data

```{r raw data}
# importing data into R using Tximport package
targets <- read_tsv("../../studyDesign.txt")# read in your study design
path <- file.path("../readMapping", targets$sample, "abundance.h5") # set file paths to your mapped data
targets <- mutate(targets, path) # add paths to your study design (only necessary for Sleuth)
Hs.anno <- useMart(biomart="ENSEMBL_MART_ENSEMBL", dataset = "hsapiens_gene_ensembl") # select 'mart' from biomaRt for annotations
Tx <- getBM(attributes=c('ensembl_transcript_id_version', # get gene symbols for each transcript ID
                         'external_gene_name'),
            mart = Hs.anno)
Tx <- as_tibble(Tx) # convert this annotation mapping file to a tibble (the tidyverse version of a dataframe)
Txi_gene <- tximport(path, #reading kallisto data into R
                     type = "kallisto", 
                     tx2gene = Tx, 
                     txOut = FALSE, 
                     countsFromAbundance = "lengthScaledTPM")
myCPM <- as_tibble(Txi_gene$abundance, rownames = "geneSymbol") # these are you counts after adjusting for transcript length
myCounts <- as_tibble(Txi_gene$counts, rownames = "geneSymbol") # these are your transcript per million (TPM) values, or counts per million (CPM) if you collapsed data to gene level

groups1 <- targets$treatment
groups1 <- factor(groups1)
sampleLabels <- targets$sample
myDGEList <- DGEList(Txi_gene$counts)
log2.cpm <- cpm(myDGEList, log=TRUE)
nsamples <- ncol(log2.cpm)
myColors <- brewer.pal(nsamples, "Paired")
log2.cpm.df <- as_tibble(log2.cpm)
colnames(log2.cpm.df) <- sampleLabels
log2.cpm.df.melt <- melt(log2.cpm.df)

ggplot(log2.cpm.df.melt, aes(x=variable, y=value, fill=variable)) +
  geom_violin(trim = FALSE, show.legend = FALSE) +
  stat_summary(fun.y = "median", geom = "point", shape = 124, size = 6, color = "black", show.legend = FALSE) +
  labs(y="log2 expression", x = "sample") +
       #title="Log2 Counts per Million (CPM)",
       #subtitle="unfiltered, non-normalized",
       #caption=paste0("produced on ", Sys.time())) +
  coord_flip() +
  theme_ipsum_rc() 
```

### Signal distribution: filtered and normalized data

```{r processed data}
cpm <- cpm(myDGEList)
keepers <- rowSums(cpm>1)>=3 #user defined
myDGEList.filtered <- myDGEList[keepers,]
myDGEList.filtered.norm <- calcNormFactors(myDGEList.filtered, method = "TMM")
log2.cpm.filtered.norm <- cpm(myDGEList.filtered.norm, log=TRUE)
log2.cpm.filtered.norm.df <- as_tibble(log2.cpm.filtered.norm)
colnames(log2.cpm.filtered.norm.df) <- sampleLabels
log2.cpm.filtered.norm.df.melt <- melt(log2.cpm.filtered.norm.df)

ggplot(log2.cpm.filtered.norm.df.melt, aes(x=variable, y=value, fill=variable)) +
  geom_violin(trim = FALSE, show.legend = FALSE) +
  stat_summary(fun.y = "median", geom = "point", shape = 124, size = 6, color = "black", show.legend = FALSE) +
  labs(y="log2 expression", x = "sample") +
       #title="Log2 Counts per Million (CPM)",
       #subtitle="filtered, TMM normalized",
       #caption=paste0("produced on ", Sys.time())) +
  coord_flip() +
  theme_ipsum_rc() 
```

### PCA plot

```{r PCA}
pca.res <- prcomp(t(log2.cpm.filtered.norm), scale.=F, retx=T)
pc.var<-pca.res$sdev^2
pc.per<-round(pc.var/sum(pc.var)*100, 1)

pca.res.df <- as_tibble(pca.res$x)
ggplot(pca.res.df, aes(x=PC1, y=PC3, color=groups1)) +
  geom_point(size=5) +
  xlab(paste0("PC1 (",pc.per[1],"%",")")) + 
  ylab(paste0("PC3 (",pc.per[3],"%",")")) + 
  theme_ipsum_rc() +
  theme(legend.title=element_blank())

```

Row {data-height=400}
------------------------------------------------------------------------------

### Filtered and normalized counts per million (CPM)

```{r filtered and normalized data}
mydata.df <- as_tibble(log2.cpm.filtered.norm, rownames = "geneSymbol")
colnames(mydata.df) <- c("geneSymbol", sampleLabels)
#user defined
mydata.df <- mutate(mydata.df,
                    uninfected.AVG = (uninf_rep1 + uninf_rep2 + uninf_rep1)/3, 
                    crypto.wt.AVG = (crypto.wt_rep1 + crypto.wt_rep2 + crypto.wt_rep3)/3,
                    crypto.mut.AVG = (crypto.mut_rep1 + crypto.mut_rep2 + crypto.mut_rep3)/3,
                    #now make columns comparing each of the averages above that you're interested in
                    LogFC.crypto.wt_vs_uninfected = (crypto.wt.AVG - uninfected.AVG),
                    LogFC.crypto.mut_vs_uninfected = (crypto.mut.AVG - uninfected.AVG)) %>%
  mutate_if(is.numeric, round, 2)

mydata.sort <- mydata.df %>%
  dplyr::arrange(desc(LogFC.crypto.wt_vs_uninfected)) %>%
  dplyr::select(geneSymbol, uninfected.AVG, crypto.wt.AVG, LogFC.crypto.wt_vs_uninfected)

datatable(mydata.df, 
          extensions = c('KeyTable', "FixedHeader"), 
          options = list(keys = TRUE, searchHighlight = TRUE, pageLength = 5, lengthMenu = c("10", "25", "50", "100"))) %>%
  formatRound(columns=c(1:6), digits=2)
```

DEGs {data-orientation=columns data-icon="fa-random"}
===================================== 

Column {data-width=500}
------------------------------------------------------------------------------

### Volcano plot

```{r volcano plot}
design <- model.matrix(~0 + groups1)
colnames(design) <- levels(groups1)
v.DEGList.filtered.norm <- voom(myDGEList.filtered.norm, design, plot = FALSE)
contrast.matrix <- makeContrasts(infection_with_WT = crypto.wt - uninfected,
                                 infection_with_Mut = crypto.mut - uninfected,
                                 levels=design)
fit <- lmFit(v.DEGList.filtered.norm, design)
fits <- contrasts.fit(fit, contrast.matrix)
ebFit <- eBayes(fits)
myTopHits <- topTable(ebFit, adjust ="BH", coef=1, number=40000, sort.by="logFC")
myTopHits <- as_tibble(myTopHits, rownames = "geneSymbol")
ggplotly(ggplot(myTopHits, aes(y=-log10(adj.P.Val), x=logFC, text = paste("Symbol:", geneSymbol))) +
           geom_point(size=2) +
           ylim(-0.5,12) +
           geom_hline(yintercept = -log10(0.01), linetype="longdash", colour="grey", size=1) +
           geom_vline(xintercept = 1, linetype="longdash", colour="#BE684D", size=1) +
           geom_vline(xintercept = -1, linetype="longdash", colour="#2C467A", size=1) +
           theme_ipsum_rc())
```


Column {.tabset data-width=500}
------------------------------------------------------------------------------

### heatmap

```{r heatmap}
#heatmap of all DEGs
#create additional heatmaps for co-expressed modules or a priori list of genes
design <- model.matrix(~0 + groups1)
colnames(design) <- levels(groups1)
v.DEGList.filtered.norm <- voom(myDGEList.filtered.norm, design, plot = FALSE)
fit <- lmFit(v.DEGList.filtered.norm, design)
contrast.matrix <- makeContrasts(infection_with_WT = crypto.wt - uninfected,
                                 infection_with_Mut = crypto.mut - uninfected,
                                 levels=design)

fits <- contrasts.fit(fit, contrast.matrix)
ebFit <- eBayes(fits)

results <- decideTests(ebFit, method="global", adjust.method="BH", p.value=0.01, lfc=2)
colnames(v.DEGList.filtered.norm$E) <- sampleLabels
diffGenes <- v.DEGList.filtered.norm$E[results[,1] !=0 | results[,2] !=0,]

myheatcolors2 <- colorRampPalette(colors=c("yellow","white","blue"))(100)
clustRows <- hclust(as.dist(1-cor(t(diffGenes), method="pearson")), method="complete") #cluster rows by pearson correlation
clustColumns <- hclust(as.dist(1-cor(diffGenes, method="spearman")), method="complete")
module.assign <- cutree(clustRows, k=2)
module.color <- rainbow(length(unique(module.assign)), start=0.1, end=0.9) 
module.color <- module.color[as.vector(module.assign)] 
d3heatmap(diffGenes,
          colors = myheatcolors2,
          Rowv=as.dendrogram(clustRows),
          row_side_colors = module.color,
          dendrogram = "none",
          scale='row'
          )
```

### DEG table

```{r DEG table}
diffGenes.df <- as_tibble(diffGenes, rownames = "geneSymbol")

diffGenes.df <- mutate(diffGenes.df,
                    uninfected.AVG = (uninf_rep1 + uninf_rep2 + uninf_rep1)/3, 
                    crypto.wt.AVG = (crypto.wt_rep1 + crypto.wt_rep2 + crypto.wt_rep3)/3,
                    crypto.mut.AVG = (crypto.mut_rep1 + crypto.mut_rep2 + crypto.mut_rep3)/3,
                    #now make columns comparing each of the averages above that you're interested in
                    LogFC = (crypto.wt.AVG - uninfected.AVG),
                    LogFC.crypto.mut_vs_uninfected = (crypto.mut.AVG - uninfected.AVG)) %>%
  mutate_if(is.numeric, round, 2)

diffGenes.sort <- diffGenes.df %>%
  dplyr::arrange(desc(LogFC)) %>%
  dplyr::select(geneSymbol, uninfected.AVG, crypto.wt.AVG, LogFC)

datatable(diffGenes.sort, 
          extensions = c('KeyTable', "FixedHeader"), 
          options = list(keys = TRUE, searchHighlight = TRUE, pageLength = 25, lengthMenu = c("10", "25", "50", "100"))) %>%
  formatRound(columns=c(1:6), digits=2)

```

About {data-orientation=rows data-icon="fa-comment-alt"}
===================================== 

Row {data-height=1000}
------------------------------------------------------------------------------

Use this page to describe the results in your own words.  Some things to think about:

* What are the key takeaways from the analysis?
* What types of analyses would you want to do next?
* Based on your analysis, what wet-lab experiments would you pursue?
* How could you expand on or otherwise enhance this data dashboard?



		</article>

	</div>

</section>

		</div>

	</div>


	<footer class="footer">

	<div class="wrap">

		<p class="footer__text">Teaching students how to use open-source tools to analyze RNAseq data since 2015.</p>

		<div class="footer__copyright">
			<span>© 2019 DIY.transcriptomics</span>
			<a href="https://jekyllthemes.io" target="_blank">Jekyll Themes</a>
		</div>

		<ul class="socials">
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
</ul>

	</div>

</footer>


	<!-- Javascript Assets -->
	<script src="https://ajax.googleapis.com/ajax/libs/jquery/3.2.1/jquery.min.js"></script>
	<script src="/js/personal-min.js"></script>

	
	<!-- Extra Footer JS Code -->
	
	


</body>

</html>