---
title: "Zomato Dashboard"
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: scroll
---

```{r setup, include=FALSE}
library(flexdashboard)
library(knitr)
library(DT)
library(scales)
library(plotly)
library(highcharter)
library(leaflet)
library(tidyverse)
library(xlsx)
library(viridisLite)
library(RColorBrewer)
library(wordcloud)
library(plotly)
library(wordcloud2)
library(rpivotTable)
library(htmlwidgets)
library(openintro)
library(countrycode)
library(leaflet.extras)
```


```{r}
rest_data<-read.csv("../input/zomato.csv")

country<-read.xlsx("../input/Country-Code.xlsx",1)
rest_data<-rest_data %>%inner_join(country,by="Country.Code")

```

Maps {data-orientation=rows }
==============================
Row {data-height=50}
--------------------------

Row {data-height=400}
------------------------------
### Restaurants by Country {.tabset .tabset-fade}

```{r}
ctry<-rest_data %>%group_by(Country)%>%summarize(cnt=n())%>%arrange(desc(cnt))
data(worldgeojson, package = "highcharter")
code <- countrycode(ctry$Country, 'country.name', 'iso3c')
ctry$iso3 <- code
data(worldgeojson, package = "highcharter")


 highchart() %>% 
                  hc_add_series_map(worldgeojson, ctry, value = "cnt", joinBy = "iso3") %>% 
                  hc_colorAxis(stops = color_stops()) %>% 
                  hc_legend(enabled = TRUE) %>%  
                  hc_mapNavigation(enabled = TRUE) %>%
                  hc_title(text = "Restaraunts by Country")  %>%
                  hc_tooltip(useHTML = TRUE, headerFormat = "",
                            pointFormat = "Country: {point.Country} Total Restaurant: {point.cnt}") %>%  hc_add_theme(hc_theme_google())


```

### Map2

```{r}
geo<-rest_data%>%select(Country, Longitude,Latitude)%>%group_by(Country)

l <- leaflet() %>% addTiles()


geo.df <- split(geo, geo$Country)

l2 <- l
names(geo.df) %>%
  purrr::walk( function(df) {
    l2 <<- l2 %>%
      addMarkers(data=geo.df[[df]],
                          lng=~Longitude, lat=~Latitude,
                          label=~Country,
                          popup=~Country,
                          group = df,
                          clusterOptions = markerClusterOptions(removeOutsideVisibleBounds = F),
                          labelOptions = labelOptions(noHide = T,
                                                       direction = 'auto'))
  })

l2 %>%
  addLayersControl(
    overlayGroups = names(geo.df),
    options = layersControlOptions(collapsed = FALSE)
  )
```

Row {data-height=400}
-----------------------------------

### Restaurants by Rating  

```{r}
rating<-rest_data%>%select(Restaurant.Name,Longitude,Latitude,Aggregate.rating, )%>%filter(!is.na(Aggregate.rating))%>%mutate(rating_levels=cut(Aggregate.rating,c(1,2,3,4,5),labels=c('0-1.9','2-2.9','3-3.9','4-5')))

col<-rest_data$Rating.color

pal <- colorFactor(rainbow(5),levels=rating$rating_levels)

l<-leaflet() %>% addProviderTiles(providers$CartoDB.DarkMatterNoLabels)

rating.df <- split(rating, rating$rating_levels)
l2 <- l
names(rating.df) %>%
  purrr::walk( function(df) {
    l2 <<- l2 %>%
      addCircleMarkers(data=rating.df[[df]],
                          lng=~Longitude, lat=~Latitude,
                          label=~paste0(Restaurant.Name," ",Aggregate.rating),
                                                    group = df,
                          stroke=FALSE, fillColor=~pal(rating_levels), fillOpacity = 1,  radius=~(as.numeric(rating.df[[df]]$rating_levels)))
  }
  )

l2 %>%
  addLayersControl(
    overlayGroups = names(rating.df),
    options = layersControlOptions(collapsed = FALSE)
  )%>%setView( median(rest_data$Longitude), median(rest_data$Latitude), zoom = 2 )
```

Restaurants {data-orientation=rows data-icon="fa-bar-chart" }
======================================================================

## Row {data-height=100}
-------------------------

## Row 1 {data-height=110 }
----------------------------------------------------------------------

### Restaurants  

```{r}
valueBox(length(unique(rest_data$Restaurant.Name)),icon = "ion-restaurant",color="red")
```

### Countries  

```{r}
valueBox(length(unique(rest_data$Country.Code)),icon="fa-globe",color="green")
```


### Cuisines  
```{r}
valueBox(length(unique(rest_data$Cuisines)),icon="fa-coffee",color="brown")
```

### Cities

```{r}
valueBox(length(unique(rest_data$City)),icon="fa-globe")
```


## Row 2 {data-height=250}
-----------------------------------------------------------------------

### Restaurants by Rating  

```{r}
rest_rate<-rest_data %>% group_by(Rating.text)%>%summarize(rcnt=n())
rest_rate %>%
plot_ly(labels=~Rating.text,values=~rcnt) %>%
  add_pie(hole=0.3) 
```

### Booking Facility

```{r}
rest_data %>%group_by(Has.Table.booking)%>%summarize(tbcnt=n())%>%
plot_ly(labels=~Has.Table.booking,values=~tbcnt) %>%
  add_pie(hole=0.3) 
```

### Online Develiery

```{r}
rest_data %>%group_by(Has.Online.delivery)%>%summarize(tbcnt=n())%>%
plot_ly(labels=~Has.Online.delivery,values=~tbcnt) %>%
  add_pie(hole=0.3)
```

## Row 3 {.tabset .tabset-fade}
----------------------------------------------------------------------


### Restaurants per Country

```{r}
ctry<-rest_data %>%group_by(Country)%>%summarize(cnt=n())%>%arrange(desc(cnt))
highchart() %>%
  hc_xAxis(categories = ctry$Country) %>% 
  hc_add_series(name = "No of Restaraunts", data = ctry$cnt, colorByPoint =  TRUE) %>% 
    hc_chart(type = "bar", options3d = list(enabled = TRUE, beta = 1, alpha = 1)) %>% hc_add_theme(hc_theme_google())
```

### Top Cities
```{r}

city<-rest_data %>%group_by(City)%>%summarize(cnt=n())%>%arrange(desc(cnt))%>%head(15)
highchart() %>%
  hc_xAxis(categories = city$City) %>% 
  hc_add_series(name = "No of Restaraunts", data = city$cnt, colorByPoint =  TRUE) %>% 
  hc_chart(type = "column", options3d = list(enabled = TRUE, beta = 1, alpha = 1)) %>% hc_add_theme(hc_theme_google())

```

### Top Restaurants by Votes

```{r}
rest<-rest_data %>%select(Restaurant.Name,Votes)%>%arrange(desc(Votes))%>%head(10)
highchart() %>%
  hc_xAxis(categories = rest$Restaurant.Name) %>% 
  hc_add_series(name = "No of Votes", data = rest$Votes, colorByPoint =  TRUE) %>% 
  hc_chart(type = "column", options3d = list(enabled = TRUE, beta = 1, alpha = 1)) %>% hc_add_theme(hc_theme_google())
```

### Popular Restaurant Names
```{r}

res_name<-rest_data %>%group_by(Restaurant.Name)%>%summarize(rescnt=n())%>%arrange(desc(rescnt))%>%head(200)
wordcloud(res_name$Restaurant.Name,res_name$rescnt, random.order=FALSE, rot.per=0.35, use.r.layout=FALSE, colors=brewer.pal(8, "Dark2"))
```

Cuisines {data-orientation=rows icon="fa-utensils" }
======================================================================

## Row {data-height=50}
-----------------------------------------------------

## Row 1 {data-height=450}
---------------------------------------------------------------------

### Cuisines worldwide  

```{r}
cuisines <- rest_data %>%
  select(Country, Cuisines) %>%
    mutate(Cuisines=strsplit(as.character(Cuisines), ", ")) %>%
  unnest(Cuisines) %>%
    group_by(Cuisines) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count))%>%head(20)


hchart(cuisines, "treemap", hcaes(x = Cuisines, value = Count, color = Count))

```

## Row 2 {data-height=450 .tabset .tabset-fade}
---------------------------------------------------------------------------

### Cuisines - USA

```{r }
cui_all<-rest_data %>%
  select(Country, Cuisines) %>%
    mutate(Cuisines=strsplit(as.character(Cuisines), ", ")) %>%
  unnest(Cuisines) %>%
    group_by(Country,Cuisines) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count))
us<-cui_all %>%filter(Country=="United States")
wordcloud(us$Cuisines,us$Count, random.order=FALSE, rot.per=0.35, use.r.layout=FALSE, colors=brewer.pal(8, "Dark2"))

```

### Cuisines - India

```{r}
india<-cui_all %>%filter(Country=="India")
wordcloud(india$Cuisines,india$Count, random.order=FALSE, rot.per=0.35, use.r.layout=FALSE, colors=brewer.pal(8, "Dark2"))

```





Price {data-orientation=rows }
===================================

## Row {data-height=50}
-------------------------------

## Row { data-height=400}
----------------------------------

### Restaurants by Price Range


```{r}
rest_data %>% group_by(Price.range)%>%
  summarise(pcnt=n())%>%
  plot_ly(x=~Price.range,
         y=~pcnt,
         fill="blue",
         type='bar')%>%
  layout(xaxis=list(title="Restaraunts by Price Range"), yaxis=list(title="Count"))
```

### Rating Vs Votes

```{r}


hc_p1<-hchart(rest_data, "scatter",hcaes(x = Votes, y = Aggregate.rating,group=1))
  hc_p1<-hc_p1 %>%hc_title(text = "Votes vs Rating") %>% 
  hc_xAxis(title = list(text = "Votes")) %>% 
  hc_yAxis(title = list(text = "Rating")) %>% 
  hc_chart(zoomType = "xy") 

hc_p1
```

## Row {data-height=400}
-------------------------------

### Costlier Restaurant  -India

```{r}
cost<-rest_data %>%filter(Country =="India") %>%select(Restaurant.Name,Average.Cost.for.two,Currency)%>%arrange(desc(Average.Cost.for.two))%>%head(10)
highchart() %>%
  hc_xAxis(categories = cost$Restaurant.Name) %>% 
  hc_add_series(name = "Average cost for two", data = cost$Average.Cost.for.two, colorByPoint =  TRUE) %>% 
  hc_title(text = "Costlier Restaurants-India")  %>%
  hc_chart(type = "bar", options3d = list(enabled = TRUE, beta = 1, alpha = 1)) %>% hc_add_theme(hc_theme_google())
```
### United States 

```{r}
cost<-rest_data %>%filter(Country =="United States") %>%select(Restaurant.Name,Average.Cost.for.two,Currency)%>%arrange(desc(Average.Cost.for.two))%>%head(10)
highchart() %>%
  hc_xAxis(categories = cost$Restaurant.Name) %>% 
  hc_add_series(name = "Average cost for two", data = cost$Average.Cost.for.two, colorByPoint =  TRUE) %>% 
  hc_title(text = "Costlier Restaurants-US")  %>%
  hc_chart(type = "bar", options3d = list(enabled = TRUE, beta = 1, alpha = 1)) %>% hc_add_theme(hc_theme_google())
```

### Average Cost for two vs Rating

```{r}


hc_p1<-hchart(rest_data, "scatter",hcaes(x = Average.Cost.for.two, y = Aggregate.rating,group=1))
  hc_p1<-hc_p1 %>%hc_title(text = "Avg Cost vs Rating") %>% 
  hc_xAxis(title = list(text = "Average Cost for two")) %>% 
  hc_yAxis(title = list(text = "Rating")) %>% 
  hc_chart(zoomType = "xy") 

hc_p1
```

Search {data-orientation=rows data-icon="fa-list" }
================================

## Row {data-height=50}
-------------------------

## Row {data-height=900 data-width=800}
----------------------------

### Find your favorite Restaurants  - Rating >3


```{r}
dt<-rest_data %>%filter(Aggregate.rating>3) %>%select(Restaurant.Name,Country,Address,Cuisines,Average.Cost.for.two,Has.Table.booking,Has.Online.delivery,Aggregate.rating,Votes)
datatable(dt,
caption="Best Restaurants",
rownames=T,
filter="top",
options = list(  pageLength = 10
))

```

About {data-orientation=rows }
=====================================================

## Row {data-height=100}
---------------------------


## Row  
----------
Kernel was about how to build interactive dashboard in R with package **flexdashboard **

The Following options has been used in this kernel

1) Multiple Pages
2) tabset in Pages
3) Multiple plot layout
4) icons
5) valueboxes

References :
https://rmarkdown.rstudio.com/flexdashboard/using.html
